Model for DES Type Ia Supernova Cosmology Analysis

### 2016-05-11 tex/toy.tex
fix likelihood discussion.

The ambition is to construct a model to determine the likelihood of DES-like
supernova data.  We start with a toy model and plan to evolve its complexity to
have the fidelity necessary for the DES analysis.
The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{../results//nodes_pgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   \label{pgm:fig}}
\end{figure}

In the toy model there are three types of data that
may be observed for each supernova:
\begin{itemize}
\item $c_o$: Counts at peak brightness, one or more independent measurements.
\item $z_o$: Redshift.
\item $T_o$: Transient type.
\end{itemize}

>>>>

The ambition is to construct a model to determine the likelihood of DES-like
supernova data. 
The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{../results//hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   \label{pgm:fig}}
\end{figure}

The model has three types of observables $O$ that
may be observed for each supernova:
\begin{itemize}
\item $c_o$: Counts at peak brightness, one or more independent measurements.
\item $z_o$: Redshift.
\item $T_o$: Transient type.
\end{itemize}

There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
host-galaxy properties, and transient properties.
There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $T$.

The likelihood is expressed as
\begin{align*}
\mathcal{L}(O| M, M_I) & = \frac{\epsilon(O, M, M_I) p(O|M, M_I)}{\int dO\, \epsilon(O, M, M_I) p(O|M, M_I)},
\end{align*}
where $p(O|M, M_I)$ is the pdf of all possible observables, and $\epsilon(O, M, M_I)$
is the probability those observables will be included in the experimental analysis.

The dependencies of the parameters are charted through the PGM to give for transients with a type measurement
\begin{align*}
p(O| M, M_I) & = P(T_o | T) p(z | M) p(c_o | L, M)
\end{align*}
and for transients without a type measurement
\begin{align*}
p(O| M, M_I) & = p(z | M) p(c_o | L, M),
\end{align*}
where the observable measurements are taken to be independent.

The efficiency $\epsilon$ depends on the properties of the experiment and the sample selection criteria:
the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, the typing
is successful, and that the type is SN~Ia:
\begin{align*}
\epsilon(O, M, M_I) & = p(SNIa, typed, sample, discovered |O, M, L, T)\\
& = p(SNIa|  typed, T_o)p(typed | sample, c_o)p(sample| discovered, c_o)p(discovered |c_o),
\end{align*}
where transient discovery, sample selection, and typing is based on observed photometry and
identification as a SN~Ia is based on observed type.
